Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-08 first_review/response-to-reviewers.tex

%Alexander, P. M., LeGrande, A. N., Fischer, E., Tedesco, M., Fettweis, X., Kelley, M., Nowicki, S. M. J., & Schmidt, G. A. (2019) Simulated Greenland surface mass balance in the GISS ModelE2 GCM: Role of the ice sheet surface. J. Geophys. Res., 124 (3), 750- 765.

\subsection*{Specific comments}

1. Lines 19-24: The language is confusing here. The quasi-uniform grids should be introduced with a brief description of how they are constructed, and what “quasi- uniform” means. Then their performance in terms of surface mass balance can be discussed, followed by the impact of grid refinement on the simulation. Also perhaps it can be mentioned here how one dycore is linked to two lon-lat configurations and another is linked to the quasi-uniform unstructured grid. \newline

{\color{blue}{We agree that this is confusing and have added a sentence describing the different grids/dycores in more detail, before talking about their SMB performance.}}.

@@

2. Line 38: Again, please explain the meaning of “quasi-uniform unstructured grid”  \newline

{\color{blue}{See above.}}.

3. Lines 52-55: Again, it would be helpful if “globally uniform unstructured grid”, “dynamical cores”, and “variable-resolution grids” can be briefly explained. Additionally, perhaps it should be explained here that two dycores are examined, one linked to the longitude latitude grid, and the other linked to unstructured grids.  \newline

{\color{blue}{Added clarification for the meaning of quasi-uniform and variable-resolution. Decided against getting into the specifics of which dycore uses which grid, as this becomes apparent later on in the introduction}}.

### 2022-09-08 arcticgrids.tex
Updates from Overleaf

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two new Arctic refined mesh configurations.
\item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
\end{keypoints}

@@

Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates. The limited area domain used by RCMs permits the use of high-resolution grids capable of resolving SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}. However, unlike GCMs, RCMs are not a freely evolving system, and the atmospheric state must be prescribed at the lateral boundaries of the model domain. The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system. In addition, the boundary conditions are derived from a separate host model, which introduces inconsistencies due to differences in model design between the host model and the RCM.

In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution. The simulated SMB compared favorably to the SMB from RCMs and observations. The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios.

@@

In CESM2.2, the SE numerics have been enhanced to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.

@@

We evaluate model simulations on six different grids in this study (Table~\ref{tbl:table1}). The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b). We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ quadrilaterial elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}c;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.

@@

\subsection{Simulated surface mass balance (SMB)}\label{sec:simSMB}

All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation from precipitation and mass loss from ablation. The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, forming ice lenses. These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, and uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB. The area fraction of each EC is derived from a high-resolution GrIS elevation dataset. The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean. Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent. Any snow above the 10 m cap contributes towards ice accumulation in the SMB, and refreezing of liquid water within the snowpack is an additional source of ice.  Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation. For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.

@@

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}b). CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids. The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.

The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets. The SMB for total water can be expressed as:
\begin{equation}
SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
\end{equation}
where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass). The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink. Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice. Note that this SMB expression is different from the internally computed SMB described in section~\ref{sec:simSMB}.

Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}b). Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid. While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term. This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask. However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the grid-mean quantities used in method~\ref{lbl:label1}.

@@

Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO. The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values refer to the average of both RACMO ensembles. Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask. Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}. For integrated precipitation, the differences between the native and common ice mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.

The coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO. The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together. Such compensating errors underlie the importance of understanding the extent to which a model is getting the right SMB for the right reasons.

@@

The bottom panel of Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots. These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt. Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}. The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects. The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.

@@

At the regional level, there is a halo of negative cloudiness bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids. This negative cloud bias contrasts with a positive cloud bias over the ice sheet interior. This anomaly pattern is attributed to deficient orographic precipitation in the coarser model grids.  With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins. As a result, excess precipitation falls in the GrIS interior, instead of being concentrated on the steep coastal margins as shown by observations \cite{P2000GPC,VETAL2018TC}. This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter. The agreement of different observational data products on this bias lends confidence in the attribution of causes. The VR grids compare better to the observations and show that orographic precipitation in Greenland is largely resolved when the horizontal resolution is increased sufficiently.

@@

As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate. We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB. That is, the simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out. This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities. We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
\appendix
\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2. These pertain mainly to the flow over orography that, for the spectral-element dynamical core, can lead to noise aligned with the element boundaries \cite{HL2018MWR}.
\subsection{Reference profiles}
Significant improvement in removing noise for flow over orography can be achieved by using reference profiles for temperature and pressure
\begin{eqnarray}
  T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}T_{ref}}\right)},
\end{eqnarray}
 \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential. The reference Exner function is
\begin{equation}
   \Pi^{(ref)}=\left( \frac{p^{(ref)}}{p_0}\right)^{\kappa}
\end{equation}
where $\kappa=\frac{R^{(d)}}{c_p^{(d)}}$. The reference surface pressure $p_0=1000hPa$ and at each model level the reference pressure $p^{(ref)}$ is computed from $p_s^{(ref)}$ and the standard hybrid coefficients
\begin{equation}
    p^{(ref)}(\eta) = A(\eta)p_0+B(\eta)p_s^{(ref)},
\end{equation}
where $A$ and $B$ are the standard hydrid coefficients (using a dry-mass generalized vertical mass coordinate $\eta$). These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
\begin{eqnarray}
   \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\nonumber \\
   && \nonumber \\
  \nabla^4_\eta T &\rightarrow& \nabla^4_\eta \left(T-T^{(ref)}\right),\label{eq:hyperT}\\ 
    \nabla^4_\eta \delta p^{(d)} &\rightarrow& \nabla^4_\eta \left(\delta p^{(d)}-\delta p^{(ref)}\right).\label{eq:hyperDP}
\end{eqnarray}
This reduces spurious transport of temperature and mass up/down-slope due to the hyperviscosity operator.

>>>>

\begin{keypoints}
\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two new Arctic refined mesh configurations.
\item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
\item The refined Arctic meshes substantially improve the surface mass balance over all low-resolution grids.
\end{keypoints}

@@

Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates. The limited area domain used by RCMs permits the use of high-resolution grids, capable of resolving SMB processes, that can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}. However, unlike GCMs, RCMs are not a freely evolving system, and the atmospheric state must be prescribed at the lateral boundaries of the model domain. The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system. In addition, the boundary conditions are derived from a separate host model, which introduces inconsistencies due to differences in model design between the host model and the RCM.

In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution. The simulated SMB compared favorably to the SMB from RCMs and observations. The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios, in the freely evolving GCM framework.

@@

@@

We evaluate model simulations on six different grids in this study (Table~\ref{tbl:table1}). The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b). We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}c;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.

@@

\subsection{Simulated surface mass balance (SMB)}

All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation from precipitation and mass loss from ablation. The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass. These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, and uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB. The area fraction of each EC is derived from a high-resolution GrIS elevation dataset. The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean. Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent. Any snow above the 10 m cap contributes towards ice accumulation in the SMB. Refreezing of liquid water within the snowpack is an additional source of ice.  Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation. For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx})b. CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids. The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.

The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets. The SMB for total water can be expressed as:
\begin{equation}
SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
\end{equation}
where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass). The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink. Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice. Note that this SMB expression is different from the internally computed SMB described in the previous section.

Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx})b. Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid. While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term. This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask. However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the grid-mean quantities used in method~\ref{lbl:label1}.

@@

@@

Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO. The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}). Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask. Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}. For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.

The coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO. The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together. Such compensating errors highlight the importance of understanding the extent to which a model is getting the right SMB for the right reasons.

@@

The bottom panel of Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots. These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt. Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}. The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells can line the east-west transects. The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.

@@

At the regional level, there is a halo of negative cloudiness bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids. This negative cloud bias contrasts with a positive cloud bias over the ice sheet interior. This anomaly pattern has been attributed to deficient orographic precipitation in the coarser model grids.  With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins. As a result, excess precipitation falls in the GrIS interior, instead of being concentrated on the steep coastal margins as shown by observations \cite{P2000GPC,VETAL2018TC}. This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter. The agreement of different observational data products on this bias lends confidence in the attribution of causes. The VR grids compare better to the observations and show that orographic precipitation in Greenland is largely resolved when the horizontal resolution is increased sufficiently.

@@

As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate. We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB. That is, the simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out. This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities. We therefore provide the the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
\appendix
\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}
Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2. These pertain mainly to the flow over orography that, for the spectral-element dynamical core, can lead to noise aligned with the element boundaries \cite{HL2018MWR}.
\subsection{Reference profiles}
Significant improvement in removing noise for flow over orography can be achieved by using reference profiles for temperature and pressure
\begin{eqnarray}
  T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}T_{ref}}\right)},
\end{eqnarray}
 \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential. The reference Exner function is
\begin{equation}
   \Pi^{(ref)}=\left( \frac{p^{(ref)}}{p_0}\right)^{\kappa}
\end{equation}
where $\kappa=\frac{R^{(d)}}{c_p^{(d)}}$. The reference surface pressure $p_0=1000hPa$ and at each model level the reference pressure $p^{(ref)}$ is computed from $p_s^{(ref)}$ and the standard hybrid coefficients
\begin{equation}
    p^{(ref)}(\eta) = A(\eta)p_0+B(\eta)p_s^{(ref)},
\end{equation}
where $A$ and $B$ are the standard hydrid coefficients (using a dry-mass generalized vertical mass coordinate $\eta$). These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
\begin{eqnarray}
   \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\nonumber \\
   && \nonumber \\
  \nabla^4_\eta T &\rightarrow& \nabla^4_\eta \left(T-T^{(ref)}\right),\label{eq:hyperT}\\ 
    \nabla^4_\eta \delta p^{(d)} &\rightarrow& \nabla^4_\eta \left(\delta p^{(d)}-\delta p^{(ref)}\right).\label{eq:hyperDP}
\end{eqnarray}
This reduces spurious transport of temperature and mass up/down-slope due to the hyperviscosity operator.
